---
title: "`r title0`"
author: "Instructor: Emily Markowitz (Emily.Markowitz@noaa.gov)"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  pdf_document:
    toc: true
    number_sections: true
    toc_depth: 2
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, 
                      error = FALSE, message = FALSE, 
                      tinytex.verbose = TRUE)

library(webshot)
library(knitr)
```

\newpage

**`r ifelse(echoTF == TRUE, "Answers to ", "") `Questions: **

# Let's explore Tidyverse!

## Think of Tidyverse as a family of packages. Which packages are loaded with {tidyverse}? What do each of these packages do? Check out: https://tidyverse.tidyverse.org/

```{r, echo = echoTF, eval = echoTF}
# You can find out by looking at the packages that are loaded when you use
library(tidyverse)

# Or, here is a screenshot from the https://tidyverse.tidyverse.org/#usage website: 
webshot::webshot("https://tidyverse.tidyverse.org/", "tidyverse_usage.png",
        cliprect = c(500, 5, 600, 505))
```

## Which package is `pivot_wider` from?

```{r, echo = echoTF, eval = echoTF}
?pivot_wider
# {tidyr} 
```

## Which package is `rename` from?

```{r, echo = echoTF, eval = echoTF}
?rename
# {dplyr} 
```

## Can you use {tidyverse} without {base} R?

```{r, echo = echoTF, eval = echoTF}
# Nope! {Tidyverse} is built on {Base} R. 
```

# Let's play with some data! 

You can view the dataset CO2 in more detail using ` View(CO2) ` and learn about it using `?CO2`. CO2 comes from the {datasets} package which should already be automatically loaded in your R. 

Some info about the CO2 dataset: "The CO2 data frame has 84 rows and 5 columns of data from an experiment on the cold tolerance of the grass species *Echinochloa crus-galli*."

```{r, echo = TRUE}
# Note: This function or data set name (in this case, data set) 
# may occur in other packages so here I am using the "::" to say I 
# specifically want the data 'CO2' from {datasets}. 
CO2<-data.frame(datasets::CO2)
head(CO2)
```

## `rename()` the "conc" column to "Concenctration mL/L" and "Treatment" column to "Condition". The new name for the conc column is not a great name (dare I say 'tidy' name?) so we'll fix that in the next question. Assign your object here as a new obect (name up to you!). 

```{r, echo = echoTF, eval = echoTF}
CO2_a<-dplyr::rename(CO2, 
            # Note that we need "" (or '') for names with spaces (and other
            # special characters) but it doesn't matter for 'one word' names
            "Concenctration mL/L" = conc, 
            Condition = Treatment)
head(CO2_a)
```

## Use the {janitor} function `clean_names` on the new CO2 data you just created in 2a. What does it do? How did {janitor} change our "Concenctration mL/L" column? 

Again, assign your object here as a new obect (name up to you!). 

```{r, echo = echoTF, eval = echoTF}
CO2_b<-janitor::clean_names(CO2_a)
head(CO2_b)

# clean_names changed the "conc" column from "Concenctration mL/L" to "concenctration_m_l_l"
```


## Use `pivot_wider` make columns of uptake (values_from) for each plant (names_from) in your new data set from question 2b. 
This is not a 'tidy' way of looking at data, but is good practice! Assign your object here as a new obect (name up to you!). 

```{r, echo = echoTF, eval = echoTF}
CO2_c<-tidyr::pivot_wider(data = CO2_b, names_from = plant, values_from = uptake)
head(CO2_c)
```

## Use `pivot_longer` to undo what you did in 2c using the data that you created in 2c. 

To see how to get the old names back, check out the `names_to` and `values_to` variable in `?pivot_longer`. This will likely incure some new rows with `NA`s, so you'll need to remove that here with `values_drop_na`. You can check if you actually got it back to origional form by seeing if the dimensions of the data.frame are the same as the origional dataset. As stated earlier, `dim(datasets::CO2)` was 84 rows and 5 columns. 

```{r, echo = echoTF, eval = echoTF}
CO2_d<-tidyr::pivot_longer(data = CO2_c, 
                  cols = c(Qn1:Mc3), #Alternatively: cols = c(4:15),
                  names_to = "plant", 
                  values_to = "uptake", 
                  values_drop_na = TRUE)

head(CO2_d)
dim(CO2_d)
```

